Lead distance under a pickoff limit in Major League Baseball:\\A sequential game model

### 2025-12-24 articles/arxiv/paper.tex
fiddle with typesetting

\title{
  Lead distance under a pickoff limit in Major League Baseball:\\A sequential game model
}
\author[1]{Scott Powers}
\author[2]{Sivaramakrishnan Ramani}
\author[3]{Jacob Hahn}
\author[2]{Andrew J. Schaefer}
\affil[1]{Department of Sport Management, Rice University}
\affil[2]{Department of Computational Applied Mathematics and Operations Research, Rice University}
\affil[3]{Department of Statistics, Rice University}

@@

      \begin{table}
        \footnotesize
        \center
        \inputtable{model_summary.tex}
        \caption{\it Summary of estimated runner outcome probability models. We estimated four different generalized (binomial) linear mixed-effects models for pickoff (PO) attempt and success and for stolen base (SB) attempt and success. For fixed effects we report the estimated effect with one standard error, and for random effects we report the estimated standard deviation term. For 2022, we encoded disengagements as always zero to reflect no pickoff limit.}
        \label{tab:model-summary}
      \end{table}

@@

      Using the transition probabilities estimated on the empirical data, we implemented the value iteration algorithm presented in Section \ref{sec:sg-model}, discretizing the runner's action space to the nearest tenth of a foot in lead distance. Table \ref{tab:lead-by-count-two-player} reports how the optimal runner policy varies by count and prior disengagements, and Table \ref{tab:lead-by-outs-two-player} reports how the optimal runner policy varies by outs and prior disengagements. This optimal lead distance policy maximizes runs scored while taking into account the rational behavior of both players, as defined by (\ref{eqn:sg_val}). For this runner policy, the pitcher is indifferent to attempting a pickoff or not.

@@

      \begin{table}[b]
        \centering
        \inputtable{lead_by_count_one_player.tex}
        \caption{
          \it The optimal lead distance (in feet) by count and prior disengagements under the single-player MDP, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
        }
        \label{tab:lead-by-count-one-player}
      \end{table}

>>>>

\title{
  Lead distance under a pickoff limit in Major League Baseball:\\A sequential game model
}
\author[1]{Scott Powers}
\author[2]{Sivaramakrishnan Ramani}
\author[3]{Jacob Hahn}
\author[2]{Andrew J. Schaefer}
\affil[1]{\small Department of Sport Management, Rice University}
\affil[2]{\small Department of Computational Applied Mathematics and Operations Research, Rice University}
\affil[3]{\small Department of Statistics, Rice University}

@@

      \begin{table}
        \small
        \center
        \inputtable{model_summary.tex}
        \caption{\it Summary of estimated runner outcome probability models. We estimated four different generalized (binomial) linear mixed-effects models for pickoff (PO) attempt and success and for stolen base (SB) attempt and success. For fixed effects we report the estimated effect with one standard error, and for random effects we report the estimated standard deviation term. For 2022, we encoded disengagements as always zero to reflect no pickoff limit.}
        \label{tab:model-summary}
      \end{table}

@@

      Using the transition probabilities estimated on the empirical data, we implemented the value iteration algorithm from Section \ref{sec:sg-model}, discretizing the runner's lead distance action space to the nearest tenth of a foot. Table \ref{tab:lead-by-count-two-player} reports how the optimal runner policy varies by count and disengagements, and Table \ref{tab:lead-by-outs-two-player} reports how the optimal runner policy varies by outs and disengagements. This runner policy maximizes runs scored while taking into account the rational behavior of both players, as defined by (\ref{eqn:sg_val}). For this runner policy, the pitcher is indifferent to attempting a pickoff or not.

@@

      \begin{table}
        \centering
        \inputtable{lead_by_count_one_player.tex}
        \caption{
          \it The optimal lead distance (in feet) by count and prior disengagements under the single-player MDP, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
        }
        \label{tab:lead-by-count-one-player}
      \end{table}
